Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-03-25 Paper/Paper.tex
prettier equation
new shorter conclusion

We use the average squared Euclidean distance as the distortion
measure between chroma patches.  Given a pattern $p_1$
composed of elements $p_1(i,j)$, and a similar pattern $p_2$, the
distance between them is:
\begin{eqnarray}
  \dist(p_1,p_2) = \sum_{i,j} \frac{(p_1(i,j) - p_2(i,j))^2}{\size(p_1)}
  \label{eq:dist}
\end{eqnarray}
We assume $p_1$ and $p_2$ have the same size.  This is enforced by the
resampling procedure described in Section \ref{sec:approach}.

@@

\section{Applications}\label{sec:exps2}
In this section we apply the result of our method to two task
not related to clustering. We try to demonstrate that the codewords
we learned capture musically meaningful patterns, at least to
contemporary pop music. This also demonstrates the utility of unsupervised 
clustering to other MIR tasks usually tackled in a supervised
framework. In the first experiment, we show that bar information is retained 
in the codewords. Thus we can discover the bar boundaries for new songs.
The second experiment shows we can discriminate artists
based on the encoding of their songs, hence we can associate artists
with specific codes.

@@

As for enhancements, we specifically did not compress our features using
Gaussian mixtures or other generative model. That being said, these methods
could be use to develop better distance measures between patterns.
The use of the Euclidean distance is arbitrary. Summarizing patches
with Gaussians, and then comparing the distance between those Gaussians,
could reduce the influence of the noise in the distance measure.

Moving on to larger scales, we could imagine merging codes together and
splitting some on two if the clusters become too small or too large.  We could
also cluster codebooks themselves, in a similar fashion as hierarchical
Gaussian mixtures \cite{Vasconcelos2001}.
%  (similar idea used for indexing \cite{Muja2009})

>>>>

We use the average squared Euclidean distance as the distortion
measure between chroma patches.  Given a pattern $p_1$
composed of elements $p_1(i,j)$, and a similar pattern $p_2$, the
distance between them is:
\begin{eqnarray}
  \dist(p_1,p_2) = \sum_{i,j} \frac{\left(p_1(i,j) - p_2(i,j)\right)^2}{\size(p_1)}
  \label{eq:dist}
\end{eqnarray}
We assume $p_1$ and $p_2$ have the same size.  This is enforced by the
resampling procedure described in Section \ref{sec:approach}.

@@

\section{Applications}\label{sec:exps2}
In this section we apply the result of our method to two tasks
unrelated to clustering. We try to demonstrate that the codewords
we learned capture musically meaningful patterns, at least to
contemporary pop music. This also demonstrates the utility of unsupervised 
clustering to other MIR tasks usually tackled in a supervised
framework. In the first experiment, we show that bar information is retained 
in the codewords. Thus we can discover the bar boundaries for new songs.
The second experiment shows we can discriminate artists
based on the encoding of their songs, hence we can associate artists
with specific codes.

@@

The main future direction is larger scales. This implies $3$ improvements.
First, gather more songs from the Echo Nest API. Second, efficiently work
with larger codebooks. Some improvements have been published for $k$-means
that avoid comparing every code to every sample. Third, the learning
algorithm must not converge too fast. We could add a filter that selects
increasingly hard example to encode, our merge many codebooks in a similar
fashion as hierarchical Gaussian mixtures \cite{Vasconcelos2001}.

%As for enhancements, we specifically did not compress our features using
%Gaussian mixtures or other generative model. That being said, these methods
%could be use to develop better distance measures between patterns.
%The use of the Euclidean distance is arbitrary. Summarizing patches
%with Gaussians, and then comparing the distance between those Gaussians,
%could reduce the influence of the noise in the distance measure.

%Moving on to larger scales, we could imagine merging codes together and
%splitting some on two if the clusters become too small or too large.  We could
%also cluster codebooks themselves, in a similar fashion as hierarchical
%Gaussian mixtures \cite{Vasconcelos2001}.
%  (similar idea used for indexing \cite{Muja2009})
